---
output:
  html_document: default
  pdf_document: default
---
```{r}
draft <- read.csv("nbaplayersdraft.csv")
```

```{r}
duke_players <- draft %>%
  filter(draft$college == "Duke" , draft$year <= "2000")
  
team_count <- duke_players %>%
  count(duke_players$team) %>%
  arrange(desc(n))

team_count
```
#The NBA teams who have drafted the most players from Duke in or before the 2000 draft are the Dallas Mavericks, Minnesota Timberwolves, and the Phoenix Suns. Each team drafted 2 players.

```{r}
d_name_players <- draft %>%
  filter(substr(draft$player, 1, 1) == "D", draft$year %% 2 == 0)

team_count <- d_name_players %>%
  count(d_name_players$team) %>%
  arrange(desc(n))

team_count
```
#The teams who have drafted the most players who have a first name that starts with D and were drafted in an even year are the Boston Celtics, the Milwaukee Bucks, and the Seattle Supersonics. Each team drafted 7 players.

```{r}
first_round_picks <- draft %>%
  filter(draft$team == "ATL") %>%
  filter(overall_pick <= 30)

ggplot(merged_data, aes(x = first_round_picks$year, y = first_round_picks$overall_pick)) +
  geom_point() +
  labs(x = "Year", y = "First Round Pick",
       title = "Relationship between Year and Overall Pick in the First Round for the Atlanta Hawks")
```

#This graph shows the First round picks for each year. You can see that in 2001 the Hawks had a high pick, meaning that in the 2000-2001 season, their record was not great. However, in the following draft they did not have a first round pick. This might indicate that their first round pick in 2001 made a good impact on the 2001-2002 season causing their record to be better and resulting in a lower pick. 

```{r}
position_stats <- draft %>%
  group_by(overall_pick) %>%
  summarize(
    average_minutes_played = sum(average_minutes_played),
    total_points = sum(points),
    total_assists = sum(assists),
    total_rebounds = sum(total_rebounds)
  ) %>%
  mutate(
    z_minutes_played = scale(average_minutes_played),
    z_total_points = scale(total_points),
    z_total_assists = scale(total_assists),
    z_total_rebounds = scale(total_rebounds)
  ) %>%
  na.omit() %>%
  mutate(
    overall_score = rowMeans(select(., starts_with("z_")), na.rm = TRUE)
  ) %>%
  arrange(desc(overall_score)) %>%
  mutate(rank = row_number()) %>%
  select(overall_pick, rank, overall_score)

position_stats
```
```{r}
position_stats <- draft %>%
  group_by(team)

team_performance <- position_stats %>%
  mutate(performance_diff = overall_pick - rank)

overperforming_teams <- filter(team_performance, performance_diff > 0) %>%
  arrange(desc(performance_diff)) %>%
  select(team, performance_diff)

underperforming_teams <- filter(team_performance, performance_diff < 0) %>%
  arrange(performance_diff) %>%
  select(team, performance_diff)

overperforming_teams
underperforming_teams
```
#Unfortunately, I could not figure out how to find over/underperforming teams based on their overall picks and the value of each draft slot found above. 